---
title: "Project 2 New"
author: "Derek, Garrett, Hud"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cerulean
    fontsize: 16pt
---


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(aplore3)
library(ggplot2)
library(plotly)
library(RColorBrewer)
library(pheatmap)
library(cluster)
library(ggcorrplot)
library(dplyr)
library(tidyr)
library(sjPlot)
library(sjmisc)
```

Data Format: A data.frame with 500 rows and 18 variables such as:

priorfrac - If the patient previously had a fracture<br>
age<br>
weight<br>
height<br>
bmi<br>
premeno<br>
momfrac<br>
armassist<br>
smoke<br>
raterisk<br>
fracscore<br>
fracture<br>
bonemed - Bone medications at enrollment (1: No, 2: Yes)<br>
bonemed_fu - Bone medications at follow-up (1: No, 2: Yes)<br>
bonetreat - Bone medications both at enrollment and follow-up (1: No, 2: Yes)<br>

```{r}
head(glow_bonemed)
```

Summary statistics for numeric variables

```{r, message=FALSE, warning=FALSE}
mysummary = glow_bonemed %>%
  select(age, weight, height, bmi, fracscore) %>%
  summarise_each(
    funs(min = min, 
    q25 = quantile(., 0.25), 
    median = median, 
    q75 = quantile(., 0.75), 
    max = max,
    mean = mean, 
    sd = sd,
    variance= var))
# reshape it using tidyr functions
clean.summary = mysummary %>% 
  gather(stat, val) %>%
  separate(stat, into = c("var", "stat"), sep = "_") %>%
  spread(stat, val) %>%
  select(var, min, max, mean, sd, variance)
print(clean.summary)
```
Summary statistics for categorical variables

```{r}
summary(glow_bonemed %>% select(priorfrac, premeno, momfrac, armassist, smoke, raterisk, fracture, bonemed, bonemed_fu, bonetreat))
```

No missing values

```{r}
colSums(is.na(glow_bonemed))
sum(is.na(glow_bonemed))


library(kableExtra)
# different way to present no missing values
# kable Extra library to make document more presentable
colSums(is.na(glow_bonemed)) %>%  
  kable("html", caption = "No missing values") %>% 
  kable_styling()
```

Age vs Weight: As weight increases the average age decreases<br>
Age vs Height: Weak correlation of as height increases age decreases<br>
Age vs BMI: As bmi increases the average age decreases<br>
Age vs fracscore: As age increases the average fracscore increases<br>

Weight vs Height: As height increases the average weight increases<br>
Weight vs BMI: As bmi increases the average weight increases<br>
Weight vs fracscore: As fracscore increases the average Weight decreases<br>

Height vs BMI: As bmi increases the average height and variance stay the same<br>
Height vs fracscore: As fracscore increases the average height stays the same though variance might decrease<br>

BMI vs fracscore: As fracscore increases the average bmi decreases<br>

```{r}
plot(glow_bonemed[, c(5:8, 14)])
```

Non of the following scatter plots show strong groupings for the fracture/no fracture categorical variable
```{r}
ggplot(glow_bonemed, aes(x = age, y = bmi, color = fracture)) +
  geom_jitter() +
  labs(title = "BMI vs Age")
```
```{r}
ggplot(glow_bonemed, aes(x = bmi, y = fracscore, color = fracture)) +
  geom_jitter() +
  labs(title = "Fracture Score vs BMI")
```

```{r}
ggplot(glow_bonemed, aes(x = fracscore, y = age, color = fracture)) +
  geom_jitter() +
  labs(title = "Age vs Fracture Score")
```

```{r}
ggplot(glow_bonemed, aes(x = weight, y = height, color = fracture)) +
  geom_jitter() +
  labs(title = "Height vs Weight")
```

Once again there doesn't seem to be strong groupings of the fracture categorical variable

```{r, message=FALSE, warning=FALSE}
fracture3dplot = plot_ly(glow_bonemed, 
  x = ~age, 
  y = ~height, 
  z = ~bmi, 
  color = ~fracture, 
  colors = c('#0C4B8E', '#BF382A')) %>% add_markers()
fracture3dplot
```

There are so little "yes" fracture results that the plot isn't very useful

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(glow_bonemed, aes(x = bmi, y = fracscore, colour = fracture)) + 
  geom_point() +
  geom_smooth(method = "loess", size = 1, span = 0.75) +
  ylim(-0.2, 1.2)
```

```{r, echo=FALSE}
ggplot(glow_bonemed, aes(x = bmi, y = age, colour = raterisk)) +
  geom_point() +
  geom_smooth(method = "loess", linewidth = 1, span = 0.75) +
  facet_wrap(~raterisk)
```

The boxplot shows that the mean fracscore seems to be slightly higher for smokers compared to non smokers

```{r}
ggplot(glow_bonemed, aes(x = smoke, y = fracscore)) +
  geom_boxplot() +
  labs(title = "Fracture Score Summary Statistics for Smokers vs Non Smokers")
```

Plot confirms there is a strong correlation between age/fracscore, bmi/weight

```{r}
corr_vars <- c("age", "weight", "height", "bmi", "fracscore")
corr_df <- glow_bonemed[, corr_vars]
corr_df <- cor(corr_df)
ggcorrplot(corr = corr_df, lab = TRUE, lab_size = 2,
  colors = c("#6D9EC1", "white", "#E46726")) +
  labs(title = "Correlation Between Variables") +
  theme(plot.title = element_text(hjust = .5),
  plot.subtitle = element_text(hjust = .5))
```

# Clustering EDA

```{r}
pheatmap(glow_bonemed[, c(5,8)], scale = "column", fontsize_row = 0.1, cluster_cols = F, legend = T, color = colorRampPalette(c("blue", "white", "red"), space = "rgb")(100))
```

```{r}
pheatmap(glow_bonemed[, 5:8], scale = "column", fontsize_row = 0.1, cluster_cols = F, legend = T, color = colorRampPalette(c("blue", "white", "red"), space = "rgb")(100))
```

```{r}
zScoreScale = scale(glow_bonemed[, 5:8])
zScoreDistance = dist(zScoreScale)
continuousVariableClustering = hclust(zScoreDistance, method = "complete")
plot(continuousVariableClustering)
```

# Modeling

Split the data into a training/testing set

```{r}
set.seed(4)
trainingIndices = sample(c(1:dim(glow_bonemed)[1]), dim(glow_bonemed)[1]*0.8)
trainingDataframe = glow_bonemed[trainingIndices,]
testingDataframe = glow_bonemed[-trainingIndices,]
```

Age is the only statistically significant continuous variable at the alpha = 0.2 level (p < 0.0001)

```{r}
model = glm(fracture ~ age + weight + height + bmi, data = glow_bonemed, family = "binomial")
summary(model)
AIC(model)

library(ResourceSelection) 
hoslem.test(model$y,fitted(model)) # shows non-significant test result which means this is a decent model fit

# get odds ratio for model
exp((model$coefficients)) 

# get confidence intervals
exp(confint(model))
```

```{r}
# trying to figure out how to use sjPlot to mimic what we did in unit12 prelive
#plot_model(model, type = "pred", terms = c("age", "smoke"))
```

```{r}
corr_vars <- c("age", "weight", "height", "bmi", "fracscore")
pc.result<-prcomp(glow_bonemed[, corr_vars],scale.=TRUE)
#Eigen Vectors
pc.result$rotation
#Eigen Values
eigenvals<-pc.result$sdev^2
eigenvals

```

```{r}
#Scree plot
par(mfrow = c(1,2))
plot(eigenvals,type = "l",
     main = "Scree Plot",
     ylab = "Eigen Values",
     xlab = "PC #")
plot(eigenvals / sum(eigenvals), 
     type = "l", main = "Scree Plot", 
     ylab = "Prop. Var. Explained", 
     xlab = "PC #", 
     ylim = c(0, 1))
cumulative.prop = cumsum(eigenvals / sum(eigenvals))
lines(cumulative.prop, lty = 2)
```


```{r}
# Loess curve for fracscore by bonetreatment group showing fracture or not
glow_bonemed$bonetreat.num <- ifelse(glow_bonemed$bonetreat == "No", 0, 1)
ggplot(glow_bonemed, aes(x = fracscore, y = bonetreat.num, color = fracture)) +
  geom_jitter()+ 
  geom_smooth(method="loess",size=1,span=1)+
  ylim(-.2,1.2) +
  xlab("Fracture Score") +
  ylab("Received bonetreatment at both iterations") +
  ggtitle("Difference in Fracture Score vs bonetreatment at both time points")
# shows that there is not an increased risk, i.e. no changes, in likelihood of fracture, whereas only receiving one or no treatments trends to increase the likelihood of a fracture as the fracture score goes up

# plot breaking down to see if there is any separation
ggplot(glow_bonemed, aes(x = fracscore, y = bonetreat, color = fracture)) +
  geom_jitter()

# in bonetreat, i.e. bone meds at both time points, in the no group, there appear to be higher fracture rates with increased fracscore, which would be predicted, i.e. if you received treatment at both times there doesn't appear to be a correlation in fracscore and breaking a bone (fracture), vs the group that did not receive both treatments appears to be a correlation with a higher likelihood correlating to likelihood of fracture
```
```{r}
# Loess curve for fracscore by physician group showing fracture or not
table(glow_bonemed$priorfrac) # show table of prior fractures

glow_bonemed$priorfrac.num <- ifelse(glow_bonemed$priorfrac == "No", 0, 1) # create numeric variable

glow_bonemed$fracture.num <- ifelse(glow_bonemed$fracture == "No", 0, 1) # create numeric variable

levels(glow_bonemed$fracture)

ggplot(glow_bonemed, aes(x = fracscore, y = fracture.num, color = priorfrac)) +
  geom_jitter()+ 
  geom_smooth(method="loess",size=1,span=1)+
  ylim(-.2,1.2)
# shows that there is not an increased risk associated with higher fracscore, i.e. no changes, in likelihood of an increased fracture if you previous had a fracture  whereas the group that has never had a fracture tends to increase the likelihood of a fracture as the fracture score goes up

# plot breaking down to see if there is any separation
ggplot(glow_bonemed, aes(x = bonetreat, y = priorfrac, color = fracture)) +
  geom_jitter()

```

```{r}
library(caret)
# plot
ggplot(glow_bonemed, aes(x = age, y = ifelse(glow_bonemed$smoke == "No", 0, 1), color = fracture)) +
  geom_jitter()+ 
  geom_smooth(method="loess",size=1,span=1)+
  ylim(-.2,1.2)




```


```{r feature selection using glmnet}
library(pROC)
set.seed(4)

#note CV and error metric are not really used here, but logLoss is reported for the final model.
# set tuning parameters using logloss
fitControl<-trainControl(method="repeatedcv",number=10,repeats=1,classProbs=TRUE, summaryFunction=mnLogLoss)

# build glmnet model

glmnet.fit<-train(fracture ~ . - sub_id,
                    data=trainingDataframe,
                    method="glmnet",
                    trControl=fitControl,
                    metric="logLoss")
coef(glmnet.fit$finalModel,glmnet.fit$finalModel$lambdaOpt)

#Getting predictions for glmnet for Complex model
glmnetfit.predprobs<-predict(glmnet.fit, trainingDataframe ,type="prob")


# glmnet ROC
glmnet.roc<-roc(response= trainingDataframe$fracture, predictor=glmnetfit.predprobs$No,levels=c("No","Yes"))


plot(glmnet.roc,col="steelblue")

# Save for later
# plot(glmnet.roc,add=T,col="steelblue")
# legend("bottomright",
#      legend=c("Simple", "Complex","GLMNET"),
#      col=c("black", "red","steelblue"),
#      lwd=4, cex =1, xpd = TRUE, horiz = FALSE)

```
\
Left out the following variables: bonetreat, bonemed, smoke, premeno, weight, age, phy_id.  
\
```{r complex model using interactions and/or polynominals NEED TO STILL DO THIS}
# Build complex model with interactions and/or polynomials

model1.2 = glm(fracture ~ age + weight + height + bmi + bonetreat + fracscore + armassist + bonetreat:fracscore, data = glow_bonemed, family = "binomial")
summary(model1.2)
AIC(model1.2)

library(ResourceSelection) 
hoslem.test(model1.2$y,fitted(model1.2)) # shows non-significant test result which means this is a decent model fit

# get odds ratio for model
exp((model1.2$coefficients)) 

# get confidence intervals
exp(confint(model1.2))
```


```{r complex model using LDA or QDA}
# skeleton code
library(caret)
fitControl<-trainControl(method="repeatedcv",number=5,repeats=1,classProbs=TRUE, summaryFunction=mnLogLoss)

set.seed(4)

#Version 1
lda.fit<-train(fracture ~ .,
               data=trainingDataframe,
               method="lda",
               trControl=fitControl,
               metric="logLoss")

#Computing predicted probabilities on the training data
predictions <- predict(lda.fit, trainingDataframe, type = "prob")[,"Yes"]

summary(predictions)

#Getting confusion matrix
threshold=0.0468
lda.preds<-factor(ifelse(predictions>threshold,"Yes","No"),levels=c("Yes","No"))
confusionMatrix(data = lda.preds, reference = trainingDataframe$fracture)




```



```{r nonparametric complex model using random forest}
library(ranger)
# set tuning parameters using logloss
fitControl<-trainControl(method="repeatedcv",number=5,repeats=1,classProbs=TRUE, savePredictions = T)

names(trainingDataframe)
randomForestModel<-train(fracture ~ . - sub_id,
                    data=trainingDataframe,
                    method="ranger",
                    trControl=fitControl,
                    preProc = c("center", "scale"))


summary(randomForestModel)
randomForestModel$results

library(MLeval)
result <- evalm(randomForestModel)

#get AUROC
result$roc
```

```{r effects plots for complex models}

# skeleton code for models using glm function


library(sjPlot)
library(sjmisc)
names(glow_bonemed)
#plot_model(glmnet.fit,type="pred", terms = c("fracscore"))
#plot_model(complex1,type="pred",terms=c("Pclass","Age[5,15,30,45]"))
#plot_model(complex1,type="pred",terms=c("Age","Sex","Pclass"))
```


```{r effects plots for complex models NEED TO STILL DO THIS}

# skeleton code for models using caret package (train) function

library(pROC)
#Predicting probabilities on the training data
glmnet.predprobs<-predict(glmnet.fit,Rose,type="prob")  #note if we were using a caret model type="raw"
glmnet.roc<-roc(response=Rose$Survived2,predictor=glmnet.predprobs$Survived,levels=c("Perished","Survived"))

plot(simple.roc)
plot(complex1.roc,print.thres="best",col="red",add=T)
plot(glmnet.roc,add=T,col="lightblue")
legend("bottomright",
       legend=c("Simple", "Complex","GLMNET"),
       col=c("black", "red","lightblue"),
       lwd=4, cex =1, xpd = TRUE, horiz = FALSE)

# plot(log.roc,print.thres="best") #This graph is nice because the x axis is plotted in terms of specificity rather than FPR
# auc(log.roc)

```



#Sources:<br>
  Hosmer, D.W., Lemeshow, S. and Sturdivant, R.X. (2013) Applied Logistic Regression, 3rd ed.,
  New York: Wiley
  
  https://cran.r-project.org/web/packages/aplore3/aplore3.pdf#page=11&zoom=100,132,90
