---
title: "glow_bonemed EDA"
author: "Derek Rogers, "
date: "2023-04-04"
output: 
  html_document:
    theme: cerulean
    fontsize: 16pt
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(aplore3)
library(ggplot2)
library(plotly)
```

Data Format:
  A data.frame with 500 rows and 18 variables such as:
  
  priorfrac - If the patient previously had a fracture
  age
  weight
  height
  bmi
  premeno
  momfrac
  armassist
  smoke
  raterisk
  fracscore
  fracture
  bonemed - Bone medications at enrollment (1: No, 2: Yes)
  bonemed_fu - Bone medications at follow-up (1: No, 2: Yes)
  bonetreat - Bone medications both at enrollment and follow-up (1: No, 2: Yes)


```{r}
head(glow_bonemed)
```

```{r}
summary(glow_bonemed)
```

From the age vs bmi and fracscore vs age scatter plots there doesn't seem to be any groupings or separations between fracture yes/no categorical response.

As age increases the average bmi decreases
There is a strong positive correlation between age and fracscore

```{r}
plot(glow_bonemed[, c(5:8, 14)])
```

```{r}
ggplot(glow_bonemed, aes(x = age, y = bmi, color = fracture)) +
  geom_jitter()
```

Fracture score increases as age increases

```{r}
ggplot(glow_bonemed, aes(x = fracscore, y = age, color = fracture)) +
  geom_jitter()
```

```{r}
ggplot(glow_bonemed, aes(x = weight, y = height, color = fracture)) +
  geom_jitter()
```

```{r}
fracture3dplot = plot_ly(glow_bonemed, 
  x = ~age, 
  y = ~height, 
  z = ~bmi, 
  color = ~fracture, 
  colors = c('#0C4B8E', '#BF382A')) %>% add_markers()

fracture3dplot
```

Clustering EDA

```{r}
library(RColorBrewer)
library(pheatmap)
library(cluster)
```

```{r}
pheatmap(glow_bonemed[, c(5,8)], scale = "column", fontsize_row = 0.1, cluster_cols = F, legend = T, color = colorRampPalette(c("blue", "white", "red"), space = "rgb")(100))
```

```{r}
pheatmap(glow_bonemed[, 5:8], scale = "column", fontsize_row = 0.1, cluster_cols = F, legend = T, color = colorRampPalette(c("blue", "white", "red"), space = "rgb")(100))
```



```{r}
zScoreScale = scale(glow_bonemed[, 5:8])

zScoreDistance = dist(zScoreScale)

continuousVariableClustering = hclust(zScoreDistance, method = "complete")

plot(continuousVariableClustering)
```

Sources:
  Hosmer, D.W., Lemeshow, S. and Sturdivant, R.X. (2013) Applied Logistic Regression, 3rd ed.,
  New York: Wiley
  
  https://cran.r-project.org/web/packages/aplore3/aplore3.pdf#page=11&zoom=100,132,90