---
title: "glow_bonemed EDA"
author: "Derek Rogers, "
date: "2023-04-04"
output: 
  html_document:
    theme: cerulean
    fontsize: 16pt
  
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(aplore3)
library(ggplot2)
library(plotly)
library(RColorBrewer)
library(pheatmap)
library(cluster)
library(ggcorrplot)
library(dplyr)
library(tidyr)
```

Data Format: A data.frame with 500 rows and 18 variables such as:

priorfrac - If the patient previously had a fracture<br>
age<br>
weight<br>
height<br>
bmi<br>
premeno<br>
momfrac<br>
armassist<br>
smoke<br>
raterisk<br>
fracscore<br>
fracture<br>
bonemed - Bone medications at enrollment (1: No, 2: Yes)<br>
bonemed_fu - Bone medications at follow-up (1: No, 2: Yes)<br>
bonetreat - Bone medications both at enrollment and follow-up (1: No, 2: Yes)<br>

```{r}
head(glow_bonemed)
```

```{r, message=FALSE, warning=FALSE}
mysummary = glow_bonemed %>%
  select(age, weight, height, bmi, fracscore) %>% # select variables to summarise
  summarise_each(funs(min = min, 
                      q25 = quantile(., 0.25), 
                      median = median, 
                      q75 = quantile(., 0.75), 
                      max = max,
                      mean = mean, 
                      sd = sd,
                      variance= var))

# reshape it using tidyr functions

clean.summary = mysummary %>% gather(stat, val) %>%
  separate(stat, into = c("var", "stat"), sep = "_") %>%
  spread(stat, val) %>%
  select(var, min, max, mean, sd,variance) # reorder columns
print(clean.summary)
```

```{r}
summary(glow_bonemed %>% select(priorfrac, premeno, momfrac, armassist, smoke, raterisk, fracture, bonemed, bonemed_fu, bonetreat))
```

Age vs Weight: As weight increases the average age decreases<br>
Age vs Height: Weak correlation of as height increases age decreases<br>
Age vs BMI: As bmi increases the average age decreases<br>
Age vs fracscore: As age increases the average fracscore increases<br>

Weight vs Height: As height increases the average weight increases<br>
Weight vs BMI: As bmi increases the average weight increases<br>
Weight vs fracscore: As fracscore increases the average Weight decreases<br>

Height vs BMI: As bmi increases the average height and variance stay the same<br>
Height vs fracscore: As fracscore increases the average height stays the same though variance might decrease<br>

BMI vs fracscore: As fracscore increases the average bmi decreases<br>

```{r}
plot(glow_bonemed[, c(5:8, 14)])
```
Non of the following scatter plots show strong groupings for the fracture/no fracture categorical variable
```{r}
ggplot(glow_bonemed, aes(x = age, y = bmi, color = fracture)) +
  geom_jitter()+
  labs(title = "BMI vs Age")
```
```{r}
ggplot(glow_bonemed, aes(x = bmi, y = fracscore, color = fracture)) +
  geom_jitter()+
  labs(title = "Fracture Score vs BMI")
```

```{r}
ggplot(glow_bonemed, aes(x = fracscore, y = age, color = fracture)) +
  geom_jitter()+
  labs(title = "Age vs Fracture Score")
```

```{r}
ggplot(glow_bonemed, aes(x = weight, y = height, color = fracture)) +
  geom_jitter()+
  labs(title = "Height vs Weight")
```

Once again there doesn't seem to be strong groupings of the fracture categorical variable

```{r, message=FALSE, warning=FALSE}
fracture3dplot = plot_ly(glow_bonemed, 
  x = ~age, 
  y = ~height, 
  z = ~bmi, 
  color = ~fracture, 
  colors = c('#0C4B8E', '#BF382A')) %>% add_markers()

fracture3dplot
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(glow_bonemed, aes(x = bmi, y = fracscore, colour = fracture)) + 
  geom_point() +
  geom_smooth(method = "loess", size = 1,span = 0.75) +
  ylim(-0.2, 1.2)
```

```{r, echo=FALSE}
ggplot(glow_bonemed, aes(x = bmi, y = age, colour = raterisk)) +
  geom_point() +
  geom_smooth(method = "loess", linewidth = 1, span = 0.75) +
  facet_wrap(~raterisk)
```

The boxplot shows that the mean fracscore seems to be slightly higher for smokers compared to non smokers

```{r}
ggplot(glow_bonemed, aes(x = smoke, y = fracscore)) +
  geom_boxplot()+
  labs(title = "Fracture Score Summary Statistics for Smokers vs Non Smokers")
```

```{r}
corr_vars <- c("age", "weight", "height", "bmi", "fracscore")

corr_df <- glow_bonemed[, corr_vars]
corr_df <- cor(corr_df)

ggcorrplot(corr = corr_df, lab = TRUE, lab_size = 2,
           colors = c("#6D9EC1", "white", "#E46726")) +
  labs(title = "Correlation Between Variables") +
  theme(plot.title = element_text(hjust = .5),
        plot.subtitle = element_text(hjust = .5))
```

# Clustering EDA

```{r}
pheatmap(glow_bonemed[, c(5,8)], scale = "column", fontsize_row = 0.1, cluster_cols = F, legend = T, color = colorRampPalette(c("blue", "white", "red"), space = "rgb")(100))
```



```{r}
pheatmap(glow_bonemed[, 5:8], scale = "column", fontsize_row = 0.1, cluster_cols = F, legend = T, color = colorRampPalette(c("blue", "white", "red"), space = "rgb")(100))
```



```{r}
zScoreScale = scale(glow_bonemed[, 5:8])

zScoreDistance = dist(zScoreScale)

continuousVariableClustering = hclust(zScoreDistance, method = "complete")

plot(continuousVariableClustering)
```

# Modeling

```{r}

```

```{r}
corr_vars <- c("age", "weight", "height", "bmi", "fracscore")
pc.result<-prcomp(glow_bonemed[, corr_vars],scale.=TRUE)

#Eigen Vectors
pc.result$rotation

#Eigen Values
eigenvals<-pc.result$sdev^2
eigenvals
```

```{r}
#Scree plot
par(mfrow=c(1,2))
plot(eigenvals,type="l",main="Scree Plot",ylab="Eigen Values",xlab="PC #")
plot(eigenvals/sum(eigenvals),type="l",main="Scree Plot",ylab="Prop. Var. Explained",xlab="PC #",ylim=c(0,1))
cumulative.prop<-cumsum(eigenvals/sum(eigenvals))
lines(cumulative.prop,lty=2)
```

#Sources:<br>
  Hosmer, D.W., Lemeshow, S. and Sturdivant, R.X. (2013) Applied Logistic Regression, 3rd ed.,
  New York: Wiley
  
  https://cran.r-project.org/web/packages/aplore3/aplore3.pdf#page=11&zoom=100,132,90