---
title: "glow_bonemed Objective 1"
author: "Joshua Hudson"
date: "2023-14-04"
output: 
  html_document:
    theme: cerulean
    fontsize: 16pt
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(aplore3)
library(ggplot2)
library(dplyr)
library(plotly)
library(ca)
library(RColorBrewer)
library(pheatmap)
library(cluster)
library(ggcorrplot)
library(dplyr)
library(tidyr)
library(sjPlot)
library(sjmisc)
library(caret)
library(vcd)
library(FactoMineR)
```

```{r}
#??glow500
```

Data Format: A data.frame with 500 rows and 18 variables such as:

priorfrac - If the patient previously had a fracture<br>
age - Age at Enrollment (years)<br>
weight - Weight at enrollment (Kilograms)<br>
height - Height at enrollment (Centimeters)<br>
bmi - Body Mass Index (Kg/m^2)<br>
premeno - Menopause before age 45 (1: No, 2: Yes)<br>
momfrac - Mother had hip fracture (1: No, 2: Yes)<br>
armassist - Arms are needed to stand from a chair (1: No, 2: Yes)<br>
smoke - Former or current smoker (1: No, 2: Yes)<br>
raterisk - Self-reported risk of fracture (1: Less than others of the same age, 2: Same as others of the same age, 3: Greater than others of the same age<br>
fracscore - Fracture Risk Score (Composite Risk Score)<br>
fracture - Any fracture in first year (1: No, 2: Yes)<br>
bonemed - Bone medications at enrollment (1: No, 2: Yes)<br>
bonemed_fu - Bone medications at follow-up (1: No, 2: Yes)<br>
bonetreat - Bone medications both at enrollment and follow-up (1: No, 2: Yes)<br>

```{r}
glow_bonemed$fracturenum = ifelse(glow_bonemed$fracture == "No", 0, 1)
head(glow_bonemed)
```


No missing values

```{r}
colSums(is.na(glow_bonemed))
sum(is.na(glow_bonemed))
```









EDA

```{r, echo=FALSE, message=FALSE, warning=FALSE}
head(glow_bonemed)
ggplot(glow_bonemed, aes(x = age, y = fracturenum, fill = smoke)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. Age") + xlab("Age") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = age, y = fracturenum, fill = bonemed)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. Age") + xlab("Age") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = age, y = fracturenum, fill = bonemed_fu)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. Age") + xlab("Age") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = age, y = fracturenum, fill = bonetreat)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. Age") + xlab("Age") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = age, y = fracturenum, fill = armassist)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. Age") + xlab("Age") + ylab("Probability of Fracture")

#Weight
ggplot(glow_bonemed, aes(x = weight, y = fracturenum, fill = smoke)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. weight") + xlab("weight") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = weight, y = fracturenum, fill = bonemed)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. weight") + xlab("weight") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = weight, y = fracturenum, fill = bonemed_fu)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. weight") + xlab("weight") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = weight, y = fracturenum, fill = bonetreat)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. weight") + xlab("weight") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = weight, y = fracturenum, fill = armassist)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. weight") + xlab("weight") + ylab("Probability of Fracture")

#height

ggplot(glow_bonemed, aes(x = height, y = fracturenum, fill = smoke)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. height") + xlab("height") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = height, y = fracturenum, fill = bonemed)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. height") + xlab("height") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = height, y = fracturenum, fill = bonemed_fu)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. height") + xlab("height") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = height, y = fracturenum, fill = bonetreat)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. height") + xlab("height") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = height, y = fracturenum, fill = armassist)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.001)) +
  geom_smooth(method = "loess", size = 1, span =1)+ggtitle("Probability of Fracture vs. height") + xlab("height") + ylab("Probability of Fracture")

```

```{r}
head(glow_bonemed)
```

Removed BMI as it is co linear with weight
Removed Fracscore as it is a form of the response

```{r}
library(ResourceSelection)
mymodel<-glm(fracture~priorfrac+age+weight+height+premeno+momfrac+armassist+smoke+raterisk+bonemed+bonemed_fu+bonetreat,data=glow_bonemed,family="binomial")
summary(mymodel)
library(car)
vif(mymodel)
library(ResourceSelection) 
hoslem.test(mymodel$y,fitted(mymodel))
confint(mymodel, level = 0.95)
```

Reduced model

```{r}
library(ResourceSelection)
mymodelr<-glm(fracture~priorfrac+age+bonemed,data=glow_bonemed,family="binomial")
summary(mymodelr)
library(car)
vif(mymodelr)
hoslem.test(mymodelr$y,fitted(mymodel))
exp(coef(summary(mymodelr)))
exp(confint(mymodelr, level = 0.95))
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
head(glow_bonemed)
ggplot(glow_bonemed, aes(x = age, y = fracturenum, fill = priorfrac)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.01)) +
  geom_smooth(method = "loess", size = 1, span =3)+ggtitle("Probability of Fracture vs. Age") + xlab("Age") + ylab("Probability of Fracture")

ggplot(glow_bonemed, aes(x = age, y = fracturenum, fill = bonemed)) + 
  geom_point(aes(color = fracture), position = position_jitter(width = 0.03, height = 0.01)) +
  geom_smooth(method = "loess", size = 1, span =3)+ggtitle("Probability of Fracture vs. Age") + xlab("Age") + ylab("Probability of Fracture")

```